---
title: "Class 11: Structural Bioinformatics 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB Database for Biomolecular Structure Data

> Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy.
    Percentage for structures solved by X-Ray = 89%
    Percentage for structured solved by Electron Microscopy = 2.5%
  
> Q2: Determine how many HIV-1 protease structures are in the current PDB?
      92.7% for proteins of HIV-1 protease structures

Download CSV File From 
```{r}
data<-read.csv("Data Export Summary.csv")
head(data)
```
Total of Different Experimental Method
```{r}
(data$Total)
```
Total Number of Entries
```{r}
sum(data$Total)
```
Finding the Proportion of Entries From Each Method
```{r}
(data$Total / sum(data$Total)) * 100
```
Finding the Proportion That Are Protein
```{r}
sum(data$Proteins) / sum(data$Total) * 100
```
Using Significant Figures
```{r}
round(sum(data$Proteins) / sum(data$Total) *100, 2)
```

## HIV-Pr Structure Analysis

Here we will read the 1HSG PDB structure and select protein component and write out a **new protein-only** PDB format file. We then do the same for the ligand (i.e. knwn drug molecule) creating a **ligand-only** PDB file.

Info:
HIV-1 protease is an enzyme that is vital for the replication of HIV. It cleaves newly formed polypeptide chains at appropriate locations so that they form function proteins. Hence, drugs that target this protein could be vital for suppressing viral replication. A handful of drugs - called HIV-1 protease inhibitors (saquinavir, ritonavir, indinavir, nelfinavir, etc.) are currently commercially available that inhibit the function of this protein, by binding in the catalytic site that typically binds the polypeptide.

Using VMD Program
> Q3: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule in this structure?
  Hydrogen atom in the water molecule has only 1 Angstrom, so we cannot see the water molecule, except the oxygen molecule.

> Q4: There is a conserved water molecule in the binding site. Can you identify this water molecule? What residue number does this water molecule have?
  The water molecule is at residue 127.

> Q5: As you have hopefully observed HIV protease is a homodimer (i.e. it is composed of twoidentical chains). With the aid of the graphic display and the sequence viewer extension can you identify secondary structure elements that are likely to only form in the dimer rather than the monomer?
  The 3rd residue of ILE on beta sheet. 

## Section 3: Introduction to Bio3D in R

Getting Started With Bio3D
```{r}
library(bio3d)
```

Reading PDB File Data into R
```{r}
# Read pdb
pdb <- read.pdb("1hsg.pdb")
print(pdb)

#atom.select()
#write.pdb()
#trim.pdb()
#Select chain A
a.inds <- atom.select(pdb, chain="A")
```

> Q6:How many amino acid residues are there in this pdb object and what are the two nonprotein residues?
  There are 198 amino acids residues and the two nonprotein residues are HOH(water) and MK1(indinavir)

Atom Select With "ligand"
```{r}
ligand <- atom.select(pdb, "ligand", value = TRUE)
write.pdb(ligand, file ="1hsg_ligand.pdb")
```

Atom Select With "protein"
```{r}
protein <- atom.select(pdb, "protein", value = TRUE)
write.pdb(protein, file = "1hsg_protein.pdb")
```

Using Attribute
```{r}
attributes(pdb)
```

To Access Individual Attributes...Use **$attribute**
```{r}
head(pdb$atom)
```

```{r}
# Print a Subset of $atom data for the first two atoms
pdb$atom[1:2, c("eleno", "elety", "x", "y", "z")]
```

```{r}
# Note that individual $atom records can also be accessed like this
pdb$atom$elety[1:2]
```

```{r}
# Which allows us to do the following
plot.bio3d(pdb$atom$b[pdb$calpha], sse=pdb, typ="l", ylab= "B-factor")
```

> Q7. What type of R object is pdb$atom?

```{r}
str(pdb$atom)
```

```{r}
# Print a summary of the coordinate data in $xyz
pdb$xyz
```

```{r}
# Examine the row and column dimensions
dim(pdb$xyz)
```

```{r}
# Print coordinates for the first two atom
pdb$xyz[1, atom2xyz(1:2)]
```

## Section 4: Atom Selection

```{r}
# Select all C-alpha atoms (return their indices)
ca.inds <- atom.select(pdb, "calpha")
```

```{r}
# Print details of the first few selected atoms
head(pdb$atom[ca.inds$atom, ])
```

```{r}
# Select xyz coordinate
head(pdb$xyz[, ca.inds$xyz])
```

## Section 5: 3D Structure Viewing in R

```{r}
# Load the Package
library("bio3d.view")
```

```{r}
# View the 3D structure
view(pdb, "overview", col = "sse")
```

