---
title: "Class 7 R Functions and Packages"
author: "Duy Tong"
date: "10/22/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Functions Revisited

Source my functions from last day

```{r}
source("http://tinyurl.com/rescale-R")
```


```{r}
rescale(1:10)
```

```{r}
rescale( c(1, 10, 5, NA, 6))
```

#is.numeric & !is.numeric functions
> ?is.numeric
> is.numeric(1:10)
[1] TRUE
> is.numeric(c(1,5,NA,10))
[1] TRUE
> is.numeric(c(1,5,NA,10,"duy"))
[1] FALSE
> !is.numeric(c(1,5,NA,10,"duy"))
[1] TRUE
> is.numeric(c(1,5,NA,10))
[1] TRUE
> !is.numeric(c(1,5,NA,10))
[1] FALSE
# ! added to is.numeric will flip false -> true / vice versa.

#```{r}
rescale2(c(1:10, "duy"))
```

#```{r}
rescale(c(1:10, "duy"))
```
Cannot run if there is a character

#Write a function to find where there are NA elements in two input vectors.

#First make some simple input where I know the answer!
```{r}
x<- c( 1, 2, NA, 3, NA)
y<- c(NA, 3, NA, 3, 4)
```
#Looked online and found the **is.na()** function
```{r}
is.na(x)
```
#and the **which()** function tells me where the TRUE values are
```{r}
which(is.na(x))
```

```{r}
which(is.na(y))
```

```{r}
is.na(x)
is.na(y)
```
#The AND function requires two input TRUE to give a TRUE output
```{r}
is.na(x) & is.na(y)
```

```{r}
sum(c(TRUE,TRUE, FALSE,TRUE))
```
#Takine the **sum()** of TRUE FALSE vector will tell me how many TRUE elements I have. (TRUE and FALSE vector cancels out). This is my working snippet!! (CODE BELOW IS THE BODY)
```{r}
sum(is.na(x) & is.na(y))
```

#Turn Into a Function
```{r}
both_na <- function(x,y) {
  sum(is.na(x) & is.na(y))
}
```
#Double Check Your Function
```{r}
both_na(x,y)
```

#eejit proofing
```{r}
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)
```

```{r}
both_na(x,y1)
```

#What will this return?
```{r}
both_na(x, y2)
```

#Make input different length...
```{r}
x2 <- c(NA,NA)
```

```{r}
both_na(x2,y2)
```

```{r}
plot(1:10, col=c("red","blue", "green"))
```

#Recycle x/y3 -> take the first NA of x add on to the next NA on y3 and keep add on to the NA on y3 -> meaning of recycled vector.
```{r}
x <-  c(NA, NA, NA)
y3 <- c( 1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```
#If like this...
```{r}
x3 <- c(1, NA, NA)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x, y3)
```

#If like this...
```{r}
x <- c(1, NA, NA, 1, NA, NA, 1)
y3 <- c(1, NA, NA, NA, NA, NA, NA)
both_na(x3, y3)
```

#Stop() Function & Add a check for when inputs x and y are not the same length. Should do this early to catch the potential problems.
```{r}
both_na2 <- function(x,y) {
  if(length(x) != length(y)) {
    stop("Inputs x and y should be the same length!")
  }
  sum(is.na(x) & is.na(y))
}
```

#Test it
#```{r}
both_na2 (x,y3)
```


#Length() Function
```{r}
length(x)
length(y3)
```

#Write a Function Practice!
#Student 1
c(100, 100, 100, 100, 100, 100, 100, 90)
#Student 2
c(100, NA, 90, 90, 90, 90, 97, 80)
```{r}
#Student 1
s1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
#Student 2
s2 <- c(100, NA, 90, 90, 90, 90, 97, 80)

mean(s1[-which.min(s1)])
mean(s2[-which.min(s2)])
```
#na.rm = TRUE means 0
```{r}
which.min(s2)
s2[-which.min(s2)]
mean(s2[-which.min(s2)], na.rm = TRUE)
```

```{r}
any(is.na(s2))
any(is.na(s1))
```


#We have our working code now turn it into a first function. #If we use warning
```{r}
grade <- function(x) {
  if(any(is.na(x))) {
    warning("Stduent is missing homework")
  }
  mean(x[-which.min(x)], na.rm = TRUE)
}
```

```{r}
grade(s1)
grade(s2)
```
#If you only did one homework and rest are NA...
```{r}
grade2 <- function(x) {
  if(any(is.na(x))) {
    warning("Stduent is missing homework")
  }
  sum(x[-which.min(x)], na.rm = TRUE /length(x)-1)
}
```

```{r}
s3 <- c(100, NA, NA, NA, NA)
grade2(s3)
```

```{r}
url <- "https://tinyurl.com/gradeinput"
hw <-read.csv(url, row.names =1)
```

```{r}
apply(hw,1, grade)
```





